---
title: "TO414 - Assignment 02"
author: "Nick Reitnour"
date: "Due Thu Sep 29th 11:59PM"
output: html_document
---

> Please start by changing the author information above. When you are done, please submit **both** the Rmd file and the html output on Canvas. Keep saving and knitting your Rmd file to make sure there are no last minute surprises


## Starting Off

We will continue with the package `nycflights13` you used in the last assignment. Recall that this package has data on all commercial flights departing the three NYC airports. We will continue our exploration of this dataset and focus on airlines this time.

### What About the Airlines?

Lets see if all airlines are equally terrible as far as flight arrival delays are concerned. For this section you will have to make sure that airline data is coded as a `factor`.

```{r}
#1. Insert code to calculate average arrival delays by airline
library("nycflights13")
airlineAverageDelay <- tapply(flights$arr_delay, flights$carrier, mean, na.rm=TRUE)
airlineAverageDelay


#2. Insert code to draw a Bar Plot of Average Arrival Delays for all the Airlines
barplot(airlineAverageDelay)

```

Which airline has the highest average delay? Which airline has the smallest average delay? Are there airlines that actually have negative average delay? Answer below:

```
#Insert answer to above question here - in plain text

The airline with the highest average delay is F9.

The airline with the smallest average delay is AS.

Yes, there are two airlines with negative average delays are AS and HA.

```

Alright - now that we know something about which airlines are better/worse as far as arrival delay is concerned, we also would want to know whether airlines actually *make up time* when on air.

```{r}
#Insert code to calculate a new variable named airgain such that airgain = (arrival delay - departure delay)
airgain <- flights$arr_delay - flights$dep_delay

#Insert code to explore airgain data and answer the question - do airlines actually gain any time when in air
mean(airgain, na.rm=TRUE)

#Insert code to calculate average airgain for all the airlines
airlineAirgain <- tapply(airgain, flights$carrier, mean, na.rm=TRUE)
airlineAirgain

#Insert code to chart average airgain for different airlines. We want to find out which airlines are good at "catching up" when in air and which airlines actually lose time further after taking off

barplot(airlineAirgain)

```

### Is There a Time Element?

Write the R code that will help you answer the questions below. **Feel free to add any needed charts to support your thought process**. 

1. If you want avoid arrival delays, which month should you fly in? 
```{r}
#Insert Code
monthDelays <- tapply(flights$arr_delay, flights$month, mean, na.rm=TRUE)
monthDelays
barplot(monthDelays)
```
```
Answer = September (month 9)
```

2. Which day (irrespective of month) is the best day to fly for minimizing arrival delay?
```{r}
#Insert Code
dayDelays <- tapply(flights$arr_delay, flights$day, mean, na.rm=TRUE)
dayDelays
barplot(dayDelays)
```
```
Answer = the 4th of each month
```

3. Assuming that you have a choice to fly in early morning (12am-8am), morning (8am-12noon), afternoon (12noon-4pm), evening (4pm-8pm), late evening (8pm-12am). Which time slot should you choose if you wish to reach your destination with minimum delay?
```{r}
#Insert Code
determineTime <- function(x) {
  if(0 <= x & x < 800) {
    timeCategory <- "Early Morning"
  } else if(800 <= x & x < 1200) {
    timeCategory <- "Morning"
  } else if(1200 <= x & x < 1600) {
    timeCategory <- "Afternoon"
  } else if(1600 <= x & x < 2000) {
    timeCategory <- "Evening"
  } else if(2000 <= x & x < 2400) {
    timeCategory <- "Late Evening"
  } else {
    timeCategory <- "NA"
  }
  timeCategory
}

categorizeTime <- sapply(flights$sched_dep_time, FUN=determineTime)

timeFrameDelays <- tapply(flights$arr_delay, categorizeTime, mean, na.rm=TRUE)
timeFrameDelays
barplot(timeFrameDelays)
```
```
Answer = Early Morning
```

### Average Speed
What is the average speed of all flights leaving NYC on year 2013? Which airline has the fastest speed of all airlines flying out of NYC airports.
```{r}
#Insert Code
averageSpeed <- mean((flights$distance/(flights$air_time/60)), na.rm = TRUE)
averageSpeed
averageSpeedByAirline <- tapply((flights$distance/(flights$air_time/60)), flights$carrier, mean, na.rm=TRUE)
averageSpeedByAirline
barplot(averageSpeedByAirline)

```
```
AverageSpeedOfAll = 394.27 mph
FastestAirline = HA
```

## Extra Credit

> Note that extra credit will be provided for excellent use of R Graphics. Do you see possibility for informative, interesting plots based on the flights data in this assignment. If yes, provide the plots below. Well crafted plots will be rewarded with extra credits.

```{r}
# Add your code for extra credit plots here. You may want to explain what you are plotting, how to interpret the plot and why is the plot interesting and/or informative.

```

> Have fun. Let me know if I can be of assistance. 
