---
title: "Machine Learning Domain Exploration"
author: "Nick Reitnour"
date: "November 20, 2016 11:59pm"
output: html_document
---

##March Madness Overview

```
The NCAA Men's Division 1 Basketball Tournament is a 68 team bracket-style tournament that is played every year to determine the college basketball champion. Due to its unpredictableness it has become an extremely popular betting event. Last year, over $9 billion was wagered on these games and that number continues to rise.

In the states where sports gambling is legal, March Madness is an extremely lucrative business with a huge problem of unpredictableness. The individuals that can get an edge are able to leverage the "Kelly Criterion" and thus maximize winnings over the long-run.

Additionally, a recent patent by Microsoft (Patent No. 0125691) has generated increased interest in sports gambling by major technology companies and what it could mean for the future of this industry.

http://www.espn.com/chalk/story/_/id/17892685/the-future-sports-betting-how-sports-betting-legalized-united-states-the-marketplace-look-like. 
```

##Preview Data
```{r}
suppressPackageStartupMessages(library(kernlab))
suppressPackageStartupMessages(library(neuralnet))
suppressPackageStartupMessages(library(gmodels))
suppressPackageStartupMessages(library(lmtest))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(aod))
suppressPackageStartupMessages(library(C50))
suppressPackageStartupMessages(library(class))
suppressPackageStartupMessages(library(neuralnet))
regularSeason <- as.data.frame(read.table("RegularSeasonDetailedResults.csv", header=T, sep=","))
head(regularSeason)
postSeason <- as.data.frame(read.table("TourneyDetailedResults.csv", header=T, sep=","))
head(postSeason)
tournamentSeeds <- as.data.frame(read.table("TourneySeeds.csv", header = T, sep=","))
head(tournamentSeeds)
```

##Organize and Clean Data
```{r}
levels(regularSeason$Wloc) <- c(0, 1, 0)
levels(postSeason$Wloc) <- c(0, 1, 0)
normalized_regularSeason <- regularSeason
normalized_regularSeason <- as.data.frame(sapply(normalized_regularSeason, as.numeric))
normalized_postSeason <- postSeason
normalized_postSeason <- as.data.frame(sapply(normalized_postSeason, as.numeric))

normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

normalized_regularSeason <- as.data.frame(lapply(normalized_regularSeason, normalize))
normalized_postSeason <- as.data.frame(lapply(normalized_postSeason, normalize))


```
##Data Key
```
Wteam = Winning team
Wscore = Points winning team scored in game
Lteam = Losing team
Lscore = Points losing team scored in game
Wloc = Binary of whether winning team was home
Numot = Number of Overtime quarters
fgm = field goals (shots) made
fga = field goals (shots) attempts
fgm3 = 3-point field goals (shots) made
fga3 = 3-point field goals (shots) attempted
ftm = free-throws made
fta = free-throws attempted
or = rebounds on offense
dr = rebounds on defense
ast = assists
to = turnovers
stl = steals
blk = blocks
pf = personal fouls
```

##Failed Attempt: Team-By-Team Linear Regression
```
At first, I thought a team-by-team approach would generate the most accurate results. In order to do this, I started by creating a massive function to further clean the data and subsequently separate it on a team-by-team basis. I successfully did this part, and then tried to predict how many points a team would score based on their match-up in March Madness. I thought that if I could predict how much each team would score, the team that more in the match-up would essentially be the winner. I hypothesized this may generate a more accurate result than just trying to generate a win probability. However, the major mistake I made in this approach was that I included every single variable, but that only diluted the potency of my model. When I tried to run a step function to uncover only the important variables, I realized that each team has different significant variables and so it was overwhelming on how to create an unifying overarching model.
```

```{r}
keyTeamStatistics <- function(teamA, teamB) {
  winningdataTeamA <- regularSeason[regularSeason$Wteam == teamA, ]
  losingdataTeamA <- regularSeason[regularSeason$Lteam == teamA, ]
  winningdataTeamB <- regularSeason[regularSeason$Wteam == teamB, ]
  losingdataTeamB <- regularSeason[regularSeason$Lteam == teamB, ]
  
  w_teamA_Points <- winningdataTeamA$Wscore
  w_teamA_Loc <- winningdataTeamA$Wloc
  w_teamA_overtime <- winningdataTeamA$Numot
  w_teamA_FieldGoalsMade <- winningdataTeamA$Wfgm
  w_teamA_FieldGoalsAttempted <- winningdataTeamA$Wfga
  w_teamA_ThreePointersMade <- winningdataTeamA$Wfgm3
  w_teamA_ThreePointersAttempted <- winningdataTeamA$Wfga3
  w_teamA_FreethrowsMade <-winningdataTeamA$Wftm
  w_teamA_FreethrowsAttempted <- winningdataTeamA$Wfta
  w_teamA_OffenseRebounds <- winningdataTeamA$Wor
  w_teamA_DefenseRebounds <- winningdataTeamA$Wdr
  w_teamA_Assists <- winningdataTeamA$Wast
  w_teamA_Turnovers <- winningdataTeamA$Wto
  w_teamA_Steals <- winningdataTeamA$Wstl
  w_teamA_Blocks <- winningdataTeamA$Wblk
  w_teamA_PersonalFouls <- winningdataTeamA$Wpf
  w_opponentA_Points <- winningdataTeamA$Lscore
  w_opponentA_FieldGoalsMade <- winningdataTeamA$Lfgm
  w_opponentA_FieldGoalsAttempted <- winningdataTeamA$Lfga
  w_opponentA_ThreePointersMade <- winningdataTeamA$Lfgm3
  w_opponentA_ThreePointersAttempted <- winningdataTeamA$Lfga3
  w_opponentA_FreethrowsMade <-winningdataTeamA$Lftm
  w_opponentA_FreethrowsAttempted <- winningdataTeamA$Lfta
  w_opponentA_OffenseRebounds <- winningdataTeamA$Lor
  w_opponentA_DefenseRebounds <- winningdataTeamA$Ldr
  w_opponentA_Assists <- winningdataTeamA$Last
  w_opponentA_Turnovers <- winningdataTeamA$Lto
  w_opponentA_Steals <- winningdataTeamA$Lstl
  w_opponentA_Blocks <- winningdataTeamA$Lblk
  w_opponentA_PersonalFouls <- winningdataTeamA$Lpf
  
  l_teamA_Points <- losingdataTeamA$Lscore
  l_teamA_Loc <- losingdataTeamA$Wloc
  l_teamA_overtime <- losingdataTeamA$Numot
  l_teamA_FieldGoalsMade <- losingdataTeamA$Lfgm
  l_teamA_FieldGoalsAttempted <- losingdataTeamA$Lfga
  l_teamA_ThreePointersMade <- losingdataTeamA$Lfgm3
  l_teamA_ThreePointersAttempted <- losingdataTeamA$Lfga3
  l_teamA_FreethrowsMade <-losingdataTeamA$Lftm
  l_teamA_FreethrowsAttempted <- losingdataTeamA$Lfta
  l_teamA_OffenseRebounds <- losingdataTeamA$Lor
  l_teamA_DefenseRebounds <- losingdataTeamA$Ldr
  l_teamA_Assists <- losingdataTeamA$Last
  l_teamA_Turnovers <- losingdataTeamA$Lto
  l_teamA_Steals <- losingdataTeamA$Lstl
  l_teamA_Blocks <- losingdataTeamA$Lblk
  l_teamA_PersonalFouls <- losingdataTeamA$Lpf
  l_opponentA_Points <- losingdataTeamA$Wscore
  l_opponentA_FieldGoalsMade <- losingdataTeamA$Wfgm
  l_opponentA_FieldGoalsAttempted <- losingdataTeamA$Wfga
  l_opponentA_ThreePointersMade <- losingdataTeamA$Wfgm3
  l_opponentA_ThreePointersAttempted <- losingdataTeamA$Wfga3
  l_opponentA_FreethrowsMade <-losingdataTeamA$Wftm
  l_opponentA_FreethrowsAttempted <- losingdataTeamA$Wfta
  l_opponentA_OffenseRebounds <- losingdataTeamA$Wor
  l_opponentA_DefenseRebounds <- losingdataTeamA$Wdr
  l_opponentA_Assists <- losingdataTeamA$Wast
  l_opponentA_Turnovers <- losingdataTeamA$Wto
  l_opponentA_Steals <- losingdataTeamA$Wstl
  l_opponentA_Blocks <- losingdataTeamA$Wblk
  l_opponentA_PersonalFouls <- losingdataTeamA$Wpf
  
  teamA_Points <- c(w_teamA_Points, l_teamA_Points)
  teamA_Points_Mean <- mean(teamA_Points)
  teamA_Loc <- c(w_teamA_Loc, l_teamA_Loc)
  teamA_Loc_Mean <- mean(teamA_Loc)
  teamA_Overtime <- c(w_teamA_overtime, l_teamA_overtime)
  teamA_Overtime_Mean <- mean(teamA_Overtime)
  teamA_FieldGoalsMade <- c(w_teamA_FieldGoalsMade, l_teamA_FieldGoalsMade)
  teamA_FieldGoalsMade_Mean <- mean(teamA_FieldGoalsMade)
  teamA_FieldGoalsAttempted <- c(w_teamA_FieldGoalsAttempted, l_teamA_FieldGoalsAttempted)
  teamA_FieldGoalsAttempted_Mean <- mean(teamA_FieldGoalsAttempted)
  teamA_ThreePointersMade <- c(w_teamA_ThreePointersMade, l_teamA_ThreePointersMade)
  teamA_ThreePointersMade_Mean <- mean(teamA_ThreePointersMade)
  teamA_ThreePointersAttempted <- c(w_teamA_ThreePointersAttempted, l_teamA_ThreePointersAttempted)
  teamA_ThreePointersAttempted_Mean <- mean(teamA_ThreePointersAttempted)
  teamA_FreethrowsMade <-c(w_teamA_FreethrowsMade, l_teamA_FreethrowsMade)
  teamA_FreethrowsMade_Mean <- mean(teamA_FreethrowsMade)
  teamA_FreethrowsAttempted <- c(w_teamA_FreethrowsAttempted, l_teamA_FreethrowsAttempted)
  teamA_FreethrowsAttempted_Mean <- mean(teamA_FreethrowsAttempted)
  teamA_OffenseRebounds <- c(w_teamA_OffenseRebounds, l_teamA_OffenseRebounds)
  teamA_OffenseRebounds_Mean <- mean(teamA_OffenseRebounds)
  teamA_DefenseRebounds <- c(w_teamA_DefenseRebounds, l_teamA_DefenseRebounds)
  teamA_DefenseRebounds_Mean <- mean(teamA_DefenseRebounds)
  teamA_Assists <- c(w_teamA_Assists, l_teamA_Assists)
  teamA_Assists_Mean <- mean(teamA_Assists)
  teamA_Turnovers <- c(w_teamA_Turnovers, l_teamA_Turnovers)
  teamA_Turnovers_Mean <- mean(teamA_Turnovers)
  teamA_Steals <- c(w_teamA_Steals, l_teamA_Steals)
  teamA_Steals_Mean <- mean(teamA_Steals)
  teamA_Blocks <- c(w_teamA_Blocks, l_teamA_Blocks)
  teamA_Blocks_Mean <- mean(teamA_Blocks)
  teamA_PersonalFouls <- c(w_teamA_PersonalFouls, l_teamA_PersonalFouls)
  teamA_PersonalFouls_Mean <- mean(teamA_PersonalFouls)
  opponentA_Points <- c(w_opponentA_Points, l_opponentA_Points)
  opponentA_FieldGoalsMade <- c(w_opponentA_FieldGoalsMade, l_opponentA_FieldGoalsMade)
  opponentA_FieldGoalsAttempted <- c(w_opponentA_FieldGoalsAttempted, l_opponentA_FieldGoalsAttempted)
  opponentA_ThreePointersMade <- c(w_opponentA_ThreePointersMade, l_opponentA_ThreePointersMade)
  opponentA_ThreePointersAttempted <- c(w_opponentA_ThreePointersAttempted, l_opponentA_ThreePointersAttempted)
  opponentA_FreethrowsMade <- c(w_opponentA_FreethrowsMade, l_opponentA_FreethrowsMade)
  opponentA_FreethrowsAttempted <- c(w_opponentA_FreethrowsAttempted, l_opponentA_FreethrowsAttempted)
  opponentA_OffenseRebounds <- c(w_opponentA_OffenseRebounds, l_opponentA_OffenseRebounds)
  opponentA_DefenseRebounds <- c(w_opponentA_DefenseRebounds, l_opponentA_DefenseRebounds)
  opponentA_Assists <- c(w_opponentA_Assists, l_opponentA_Assists)
  opponentA_Turnovers <- c(w_opponentA_Turnovers, l_opponentA_Turnovers)
  opponentA_Steals <- c(w_opponentA_Steals, l_opponentA_Steals)
  opponentA_Blocks <- c(w_opponentA_Blocks, l_opponentA_Blocks)
  opponentA_PersonalFouls <- c(w_opponentA_PersonalFouls, l_opponentA_PersonalFouls)
  
  w_teamB_Points <- winningdataTeamB$Wscore
  w_teamB_Loc <- winningdataTeamB$Wloc
  w_teamB_overtime <- winningdataTeamB$Numot
  w_teamB_FieldGoalsMade <- winningdataTeamB$Wfgm
  w_teamB_FieldGoalsAttempted <- winningdataTeamB$Wfga
  w_teamB_ThreePointersMade <- winningdataTeamB$Wfgm3
  w_teamB_ThreePointersAttempted <- winningdataTeamB$Wfga3
  w_teamB_FreethrowsMade <-winningdataTeamB$Wftm
  w_teamB_FreethrowsAttempted <- winningdataTeamB$Wfta
  w_teamB_OffenseRebounds <- winningdataTeamB$Wor
  w_teamB_DefenseRebounds <- winningdataTeamB$Wdr
  w_teamB_Assists <- winningdataTeamB$Wast
  w_teamB_Turnovers <- winningdataTeamB$Wto
  w_teamB_Steals <- winningdataTeamB$Wstl
  w_teamB_Blocks <- winningdataTeamB$Wblk
  w_teamB_PersonalFouls <- winningdataTeamB$Wpf
  w_opponentB_Points <- winningdataTeamB$Lscore
  w_opponentB_FieldGoalsMade <- winningdataTeamB$Lfgm
  w_opponentB_FieldGoalsAttempted <- winningdataTeamB$Lfga
  w_opponentB_ThreePointersMade <- winningdataTeamB$Lfgm3
  w_opponentB_ThreePointersAttempted <- winningdataTeamB$Lfga3
  w_opponentB_FreethrowsMade <-winningdataTeamB$Lftm
  w_opponentB_FreethrowsAttempted <- winningdataTeamB$Lfta
  w_opponentB_OffenseRebounds <- winningdataTeamB$Lor
  w_opponentB_DefenseRebounds <- winningdataTeamB$Ldr
  w_opponentB_Assists <- winningdataTeamB$Last
  w_opponentB_Turnovers <- winningdataTeamB$Lto
  w_opponentB_Steals <- winningdataTeamB$Lstl
  w_opponentB_Blocks <- winningdataTeamB$Lblk
  w_opponentB_PersonalFouls <- winningdataTeamB$Lpf
  
  l_teamB_Points <- losingdataTeamB$Lscore
  l_teamB_Loc <- losingdataTeamB$Wloc
  l_teamB_overtime <- losingdataTeamB$Numot
  l_teamB_FieldGoalsMade <- losingdataTeamB$Lfgm
  l_teamB_FieldGoalsAttempted <- losingdataTeamB$Lfga
  l_teamB_ThreePointersMade <- losingdataTeamB$Lfgm3
  l_teamB_ThreePointersAttempted <- losingdataTeamB$Lfga3
  l_teamB_FreethrowsMade <-losingdataTeamB$Lftm
  l_teamB_FreethrowsAttempted <- losingdataTeamB$Lfta
  l_teamB_OffenseRebounds <- losingdataTeamB$Lor
  l_teamB_DefenseRebounds <- losingdataTeamB$Ldr
  l_teamB_Assists <- losingdataTeamB$Last
  l_teamB_Turnovers <- losingdataTeamB$Lto
  l_teamB_Steals <- losingdataTeamB$Lstl
  l_teamB_Blocks <- losingdataTeamB$Lblk
  l_teamB_PersonalFouls <- losingdataTeamB$Lpf
  l_opponentB_Points <- losingdataTeamB$Wscore
  l_opponentB_FieldGoalsMade <- losingdataTeamB$Wfgm
  l_opponentB_FieldGoalsAttempted <- losingdataTeamB$Wfga
  l_opponentB_ThreePointersMade <- losingdataTeamB$Wfgm3
  l_opponentB_ThreePointersAttempted <- losingdataTeamB$Wfga3
  l_opponentB_FreethrowsMade <-losingdataTeamB$Wftm
  l_opponentB_FreethrowsAttempted <- losingdataTeamB$Wfta
  l_opponentB_OffenseRebounds <- losingdataTeamB$Wor
  l_opponentB_DefenseRebounds <- losingdataTeamB$Wdr
  l_opponentB_Assists <- losingdataTeamB$Wast
  l_opponentB_Turnovers <- losingdataTeamB$Wto
  l_opponentB_Steals <- losingdataTeamB$Wstl
  l_opponentB_Blocks <- losingdataTeamB$Wblk
  l_opponentB_PersonalFouls <- losingdataTeamB$Wpf
  
  teamB_Points <- c(w_teamB_Points, l_teamB_Points)
  teamB_Points_Mean <- mean(teamB_Points)
  teamB_Loc <- c(w_teamB_Loc, l_teamB_Loc)
  teamB_Loc_Mean <- mean(teamB_Loc)
  teamB_Overtime <- c(w_teamB_overtime, l_teamB_overtime)
  teamB_Overtime_Mean <- mean(teamB_Overtime)
  teamB_FieldGoalsMade <- c(w_teamB_FieldGoalsMade, l_teamB_FieldGoalsMade)
  teamB_FieldGoalsMade_Mean <- mean(teamB_FieldGoalsMade)
  teamB_FieldGoalsAttempted <- c(w_teamB_FieldGoalsAttempted, l_teamB_FieldGoalsAttempted)
  teamB_FieldGoalsAttempted_Mean <- mean(teamB_FieldGoalsAttempted)
  teamB_ThreePointersMade <- c(w_teamB_ThreePointersMade, l_teamB_ThreePointersMade)
  teamB_ThreePointersMade_Mean <- mean(teamB_ThreePointersMade)
  teamB_ThreePointersAttempted <- c(w_teamB_ThreePointersAttempted, l_teamB_ThreePointersAttempted)
  teamB_ThreePointersAttempted_Mean <- mean(teamB_ThreePointersAttempted)
  teamB_FreethrowsMade <-c(w_teamB_FreethrowsMade, l_teamB_FreethrowsMade)
  teamB_FreethrowsMade_Mean <- mean(teamB_FreethrowsMade)
  teamB_FreethrowsAttempted <- c(w_teamB_FreethrowsAttempted, l_teamB_FreethrowsAttempted)
  teamB_FreethrowsAttempted_Mean <- mean(teamB_FreethrowsAttempted)
  teamB_OffenseRebounds <- c(w_teamB_OffenseRebounds, l_teamB_OffenseRebounds)
  teamB_OffenseRebounds_Mean <- mean(teamB_OffenseRebounds)
  teamB_DefenseRebounds <- c(w_teamB_DefenseRebounds, l_teamB_DefenseRebounds)
  teamB_DefenseRebounds_Mean <- mean(teamB_DefenseRebounds)
  teamB_Assists <- c(w_teamB_Assists, l_teamB_Assists)
  teamB_Assists_Mean <- mean(teamB_Assists)
  teamB_Turnovers <- c(w_teamB_Turnovers, l_teamB_Turnovers)
  teamB_Turnovers_Mean <- mean(teamB_Turnovers)
  teamB_Steals <- c(w_teamB_Steals, l_teamB_Steals)
  teamB_Steals_Mean <- mean(teamB_Steals)
  teamB_Blocks <- c(w_teamB_Blocks, l_teamB_Blocks)
  teamB_Blocks_Mean <- mean(teamB_Blocks)
  teamB_PersonalFouls <- c(w_teamB_PersonalFouls, l_teamB_PersonalFouls)
  teamB_PersonalFouls_Mean <- mean(teamB_PersonalFouls)
  opponentB_Points <- c(w_opponentB_Points, l_opponentB_Points)
  opponentB_FieldGoalsMade <- c(w_opponentB_FieldGoalsMade, l_opponentB_FieldGoalsMade)
  opponentB_FieldGoalsAttempted <- c(w_opponentB_FieldGoalsAttempted, l_opponentB_FieldGoalsAttempted)
  opponentB_ThreePointersMade <- c(w_opponentB_ThreePointersMade, l_opponentB_ThreePointersMade)
  opponentB_ThreePointersAttempted <- c(w_opponentB_ThreePointersAttempted, l_opponentB_ThreePointersAttempted)
  opponentB_FreethrowsMade <- c(w_opponentB_FreethrowsMade, l_opponentB_FreethrowsMade)
  opponentB_FreethrowsAttempted <- c(w_opponentB_FreethrowsAttempted, l_opponentB_FreethrowsAttempted)
  opponentB_OffenseRebounds <- c(w_opponentB_OffenseRebounds, l_opponentB_OffenseRebounds)
  opponentB_DefenseRebounds <- c(w_opponentB_DefenseRebounds, l_opponentB_DefenseRebounds)
  opponentB_Assists <- c(w_opponentB_Assists, l_opponentB_Assists)
  opponentB_Turnovers <- c(w_opponentB_Turnovers, l_opponentB_Turnovers)
  opponentB_Steals <- c(w_opponentB_Steals, l_opponentB_Steals)
  opponentB_Blocks <- c(w_opponentB_Blocks, l_opponentB_Blocks)
  opponentB_PersonalFouls <- c(w_opponentB_PersonalFouls, l_opponentB_PersonalFouls)
  
  
  teamA_LinearModel <- lm(teamA_Points ~ teamA_FieldGoalsMade + teamA_FieldGoalsAttempted + teamA_ThreePointersMade + teamA_ThreePointersAttempted + teamA_FreethrowsMade + teamA_FreethrowsAttempted + teamA_OffenseRebounds + teamA_DefenseRebounds + teamA_Assists + teamA_Turnovers + teamA_Steals + teamA_Blocks + teamA_PersonalFouls + opponentA_Points + opponentA_FieldGoalsMade + opponentA_FieldGoalsAttempted + opponentA_ThreePointersMade + opponentA_ThreePointersAttempted + opponentA_FreethrowsMade + opponentA_FreethrowsAttempted + opponentA_OffenseRebounds + opponentA_DefenseRebounds + opponentA_Assists + opponentA_Turnovers + opponentA_Steals + opponentA_Blocks + opponentA_PersonalFouls)
  
  teamA_SpecificValues <- data.frame(teamA_FieldGoalsMade = teamA_FieldGoalsMade_Mean, teamA_FieldGoalsAttempted = teamA_FieldGoalsAttempted_Mean, teamA_ThreePointersMade = teamA_ThreePointersMade_Mean, teamA_ThreePointersAttempted = teamA_ThreePointersAttempted_Mean, teamA_FreethrowsMade = teamA_FreethrowsMade_Mean, teamA_FreethrowsAttempted = teamA_FreethrowsAttempted_Mean, teamA_OffenseRebounds = teamA_OffenseRebounds_Mean, teamA_DefenseRebounds = teamA_DefenseRebounds_Mean, teamA_Assists = teamA_Assists_Mean, teamA_Turnovers = teamA_Turnovers_Mean, teamA_Steals = teamA_Steals_Mean, teamA_Blocks = teamA_Blocks_Mean, teamA_PersonalFouls = teamA_PersonalFouls_Mean, opponentA_Points = teamB_Points_Mean, opponentA_FieldGoalsMade = teamB_FieldGoalsMade_Mean, opponentA_FieldGoalsAttempted = teamB_FieldGoalsAttempted_Mean, opponentA_ThreePointersMade = teamB_ThreePointersMade_Mean, opponentA_ThreePointersAttempted = teamB_ThreePointersAttempted_Mean, opponentA_FreethrowsMade = teamB_FreethrowsMade_Mean, opponentA_FreethrowsAttempted = teamB_FreethrowsAttempted_Mean, opponentA_OffenseRebounds = teamB_OffenseRebounds_Mean, opponentA_DefenseRebounds = teamB_DefenseRebounds_Mean, opponentA_Assists = teamB_Assists_Mean, opponentA_Turnovers = teamB_Turnovers_Mean, opponentA_Steals = teamB_Steals_Mean, opponentA_Blocks = teamB_Blocks_Mean, opponentA_PersonalFouls = teamB_PersonalFouls_Mean)
  abc <- predict(teamA_LinearModel, newdata=teamA_SpecificValues)
  
  teamB_LinearModel <- lm(teamB_Points ~ teamB_FieldGoalsMade + teamB_FieldGoalsAttempted + teamB_ThreePointersMade + teamB_ThreePointersAttempted + teamB_FreethrowsMade + teamB_FreethrowsAttempted + teamB_OffenseRebounds + teamB_DefenseRebounds + teamB_Assists + teamB_Turnovers + teamB_Steals + teamB_Blocks + teamB_PersonalFouls + opponentB_Points + opponentB_FieldGoalsMade + opponentB_FieldGoalsAttempted + opponentB_ThreePointersMade + opponentB_ThreePointersAttempted + opponentB_FreethrowsMade + opponentB_FreethrowsAttempted + opponentB_OffenseRebounds + opponentB_DefenseRebounds + opponentB_Assists + opponentB_Turnovers + opponentB_Steals + opponentB_Blocks + opponentB_PersonalFouls)
  aaa <- teamA_Points_Mean - abc
  return(aaa)
}

#keyTeamStatistics(1101, 1103)

score <- function(teamA) {
  winningdataTeamA <- regularSeason[regularSeason$Wteam == teamA, ]
  losingdataTeamA <- regularSeason[regularSeason$Lteam == teamA, ]
  allDataTeamA <- rbind(winningdataTeamA, losingdataTeamA)
  w_teamA <- winningdataTeamA$Wscore
  l_teamA <- losingdataTeamA$Lscore
  teamA_Points <- c(w_teamA, l_teamA)
  return(teamA_Points)
}

fieldGoalsMade <- function(teamA) {
  winningdataTeamA <- regularSeason[regularSeason$Wteam == teamA, ]
  losingdataTeamA <- regularSeason[regularSeason$Lteam == teamA, ]
  allDataTeamA <- rbind(winningdataTeamA, losingdataTeamA)
  w_teamA <- winningdataTeamA$Wfgm
  l_teamA <- losingdataTeamA$Lfgm
  teamA_fieldGaolsMade <- c(w_teamA, l_teamA)
  return(teamA_Points)
}
```
##Cleaner Data
```
I think I was really struggling before because there were two separate outcomes (win and loss) combined in each row. In order to made it more suitable for my level of comprehension with R, I decided I should break everything into separate lines. The winning team gets one row all to itself and the losing team gets one row all to itself.
```
```{r}
winningteamData <- regularSeason
names(winningteamData)[2] <- "Win"
winningteamData$Win <- 1
names(winningteamData)[3] <- "Team"
names(winningteamData)[4] <- "TeamScore"
names(winningteamData)[5] <- "Opp"
names(winningteamData)[6] <- "OppScore"
names(winningteamData)[7] <- "Home"
names(winningteamData)[9] <- "TeamFGM"
names(winningteamData)[10] <- "TeamFGA"
names(winningteamData)[11] <- "TeamFGM3"
names(winningteamData)[12] <- "TeamFGA3"
names(winningteamData)[13] <- "TeamFTM"
names(winningteamData)[14] <- "TeamFTA"
names(winningteamData)[15] <- "TeamOR"
names(winningteamData)[16] <- "TeamDR"
names(winningteamData)[17] <- "TeamAST"
names(winningteamData)[18] <- "TeamTO"
names(winningteamData)[19] <- "TeamSTL"
names(winningteamData)[20] <- "TeamBLK"
names(winningteamData)[21] <- "TeamPF"
names(winningteamData)[22] <- "OppFGM"
names(winningteamData)[23] <- "OppFGA"
names(winningteamData)[24] <- "OppFGM3"
names(winningteamData)[25] <- "OppFGA3"
names(winningteamData)[26] <- "OppFTM"
names(winningteamData)[27] <- "OppFTA"
names(winningteamData)[28] <- "OppOR"
names(winningteamData)[29] <- "OppDR"
names(winningteamData)[30] <- "OppAST"
names(winningteamData)[31] <- "OppTO"
names(winningteamData)[32] <- "OppSTL"
names(winningteamData)[33] <- "OppBLK"
names(winningteamData)[34] <- "OppPF"
winningteamData[7] <- NULL

losingteamData <- regularSeason
names(losingteamData)[2] <- "Win"
losingteamData$Win <- 0
names(losingteamData)[3] <- "Opp"
names(losingteamData)[4] <- "OppScore"
names(losingteamData)[5] <- "Team"
names(losingteamData)[6] <- "TeamScore"
names(losingteamData)[7] <- "Home"
names(losingteamData)[9] <- "OppFGM"
names(losingteamData)[10] <- "OppFGA"
names(losingteamData)[11] <- "OppFGM3"
names(losingteamData)[12] <- "OppFGA3"
names(losingteamData)[13] <- "OppFTM"
names(losingteamData)[14] <- "OppFTA"
names(losingteamData)[15] <- "OppOR"
names(losingteamData)[16] <- "OppDR"
names(losingteamData)[17] <- "OppAST"
names(losingteamData)[18] <- "OppTO"
names(losingteamData)[19] <- "OppSTL"
names(losingteamData)[20] <- "OppBLK"
names(losingteamData)[21] <- "OppPF"
names(losingteamData)[22] <- "TeamFGM"
names(losingteamData)[23] <- "TeamFGA"
names(losingteamData)[24] <- "TeamFGM3"
names(losingteamData)[25] <- "TeamFGA3"
names(losingteamData)[26] <- "TeamFTM"
names(losingteamData)[27] <- "TeamFTA"
names(losingteamData)[28] <- "TeamOR"
names(losingteamData)[29] <- "TeamDR"
names(losingteamData)[30] <- "TeamAST"
names(losingteamData)[31] <- "TeamTO"
names(losingteamData)[32] <- "TeamSTL"
names(losingteamData)[33] <- "TeamBLK"
names(losingteamData)[34] <- "TeamPF"
losingteamData[7] <- NULL

singleRowData_All <- rbind(winningteamData, losingteamData)


winningteamData_PostSeason <- postSeason
names(winningteamData_PostSeason)[2] <- "Win"
winningteamData_PostSeason$Win <- 1
names(winningteamData_PostSeason)[3] <- "Team"
names(winningteamData_PostSeason)[4] <- "TeamScore"
names(winningteamData_PostSeason)[5] <- "Opp"
names(winningteamData_PostSeason)[6] <- "OppScore"
names(winningteamData_PostSeason)[7] <- "Home"
names(winningteamData_PostSeason)[9] <- "TeamFGM"
names(winningteamData_PostSeason)[10] <- "TeamFGA"
names(winningteamData_PostSeason)[11] <- "TeamFGM3"
names(winningteamData_PostSeason)[12] <- "TeamFGA3"
names(winningteamData_PostSeason)[13] <- "TeamFTM"
names(winningteamData_PostSeason)[14] <- "TeamFTA"
names(winningteamData_PostSeason)[15] <- "TeamOR"
names(winningteamData_PostSeason)[16] <- "TeamDR"
names(winningteamData_PostSeason)[17] <- "TeamAST"
names(winningteamData_PostSeason)[18] <- "TeamTO"
names(winningteamData_PostSeason)[19] <- "TeamSTL"
names(winningteamData_PostSeason)[20] <- "TeamBLK"
names(winningteamData_PostSeason)[21] <- "TeamPF"
names(winningteamData_PostSeason)[22] <- "OppFGM"
names(winningteamData_PostSeason)[23] <- "OppFGA"
names(winningteamData_PostSeason)[24] <- "OppFGM3"
names(winningteamData_PostSeason)[25] <- "OppFGA3"
names(winningteamData_PostSeason)[26] <- "OppFTM"
names(winningteamData_PostSeason)[27] <- "OppFTA"
names(winningteamData_PostSeason)[28] <- "OppOR"
names(winningteamData_PostSeason)[29] <- "OppDR"
names(winningteamData_PostSeason)[30] <- "OppAST"
names(winningteamData_PostSeason)[31] <- "OppTO"
names(winningteamData_PostSeason)[32] <- "OppSTL"
names(winningteamData_PostSeason)[33] <- "OppBLK"
names(winningteamData_PostSeason)[34] <- "OppPF"
winningteamData_PostSeason[7] <- NULL

losingteamData_PostSeason <- postSeason
names(losingteamData_PostSeason)[2] <- "Win"
losingteamData_PostSeason$Win <- 0
names(losingteamData_PostSeason)[3] <- "Opp"
names(losingteamData_PostSeason)[4] <- "OppScore"
names(losingteamData_PostSeason)[5] <- "Team"
names(losingteamData_PostSeason)[6] <- "TeamScore"
names(losingteamData_PostSeason)[7] <- "Home"
names(losingteamData_PostSeason)[9] <- "OppFGM"
names(losingteamData_PostSeason)[10] <- "OppFGA"
names(losingteamData_PostSeason)[11] <- "OppFGM3"
names(losingteamData_PostSeason)[12] <- "OppFGA3"
names(losingteamData_PostSeason)[13] <- "OppFTM"
names(losingteamData_PostSeason)[14] <- "OppFTA"
names(losingteamData_PostSeason)[15] <- "OppOR"
names(losingteamData_PostSeason)[16] <- "OppDR"
names(losingteamData_PostSeason)[17] <- "OppAST"
names(losingteamData_PostSeason)[18] <- "OppTO"
names(losingteamData_PostSeason)[19] <- "OppSTL"
names(losingteamData_PostSeason)[20] <- "OppBLK"
names(losingteamData_PostSeason)[21] <- "OppPF"
names(losingteamData_PostSeason)[22] <- "TeamFGM"
names(losingteamData_PostSeason)[23] <- "TeamFGA"
names(losingteamData_PostSeason)[24] <- "TeamFGM3"
names(losingteamData_PostSeason)[25] <- "TeamFGA3"
names(losingteamData_PostSeason)[26] <- "TeamFTM"
names(losingteamData_PostSeason)[27] <- "TeamFTA"
names(losingteamData_PostSeason)[28] <- "TeamOR"
names(losingteamData_PostSeason)[29] <- "TeamDR"
names(losingteamData_PostSeason)[30] <- "TeamAST"
names(losingteamData_PostSeason)[31] <- "TeamTO"
names(losingteamData_PostSeason)[32] <- "TeamSTL"
names(losingteamData_PostSeason)[33] <- "TeamBLK"
names(losingteamData_PostSeason)[34] <- "TeamPF"
losingteamData_PostSeason[7] <- NULL

singleRowData_All_PostSeason <- rbind(winningteamData_PostSeason, losingteamData_PostSeason)

```
##Is this the year?
```
In the 2012-2013 Season, Michigan Basketball made it to the NCAA Men's Basketball National Championship game. However, fast forward a few years and in the 2014-2015 Season and the Michigan Wolverines didn't even make the tournament. Every NCAA basketball season is unique and so I hypothesize that it would be smart to see if I can make a universal model for a particular season that can predict all winners. Making one for all 300+ teams could be a little too much ...

For simplicity's sake, I am just going to take the NCAA Men's Basketball Data for 2015. 
```
```{r}
singleRowData_All2015 <- singleRowData_All[singleRowData_All$Season == 2015, ]
tail(singleRowData_All_PostSeason)
singleRowData_All2015_PostSeason <- singleRowData_All_PostSeason[singleRowData_All_PostSeason$Season == 2015, ]
head(singleRowData_All2015_PostSeason)
```
###Regression Models
```{r}
singleRowData_All2015_Regression <- singleRowData_All2015
singleRowData_All2015_Regression[1] <- NULL #Remove 'Season'
singleRowData_All2015_Regression[2] <- NULL #Remove 'Team'
singleRowData_All2015_Regression[3] <- NULL #Remove 'Opponent'
singleRowData_All2015_Regression[4] <- NULL #Remove 'Numot'
singleRowData_All2015_Regression[8] <- NULL #Remove 'TeamFTM' because causing problems with regression
singleRowData_All2015_Regression[20] <- NULL #Remove 'OppFTM' because causing problems with regression
```
####Linear Regression Model
```{r}
linear_Model <- lm(Win ~ ., data = singleRowData_All2015_Regression)
stepLinearModel <- step(object = linear_Model, direction = "backward", trace=0)
summary(stepLinearModel)
```
```
I noticed two things when I did this. 

First was that because I broke up the Winning team and the Losing team into two separate rows, it resulted in redundancy in my data. The "Opp" columns were essentially the same number but of the opposite sign of the "Team" data. Therefore I should delete these columns.

Secondly, I realized that I shouldn't be running a Linear Regression, but rather a Binomial Logit Model to predict whether the team will win or not.
```
####Logit Regression Model
```{r}
singleRowData_All2015_Regression_Logit <- singleRowData_All2015_Regression
singleRowData_All2015_Regression_Logit[3] <- NULL
singleRowData_All2015_Regression_Logit[15:ncol(singleRowData_All2015_Regression_Logit)] <- NULL
logit_Model <- glm(Win ~ ., data= singleRowData_All2015_Regression_Logit, family= "binomial")
summary(logit_Model)
```
```
I got a much cleaner result when it came to the Logit Regression.
```
###Decision Tree Model
```
While the Logit Model was good, it was a very old concept that I learned at the start of the semester. As a result, I decided to go to the other extreme and see how a Decision Tree would do.
```
```{r}
singleRowData_All2015_DecisionTree <- singleRowData_All2015
singleRowData_All2015_DecisionTree[1] <- NULL #Remove 'Season' since given
singleRowData_All2015_DecisionTree[3] <- NULL #Remove 'Score'
singleRowData_All2015_DecisionTree[4] <- NULL #Remove 'Score'
singleRowData_All2015_DecisionTree$Win <- as.factor(singleRowData_All2015_DecisionTree$Win)
decisionTree_Model <- C5.0(singleRowData_All2015_DecisionTree[-1], singleRowData_All2015_DecisionTree$Win)
summary(decisionTree_Model)
```
###KNN Model
```{r}
regularSeason_2015_KNN <- singleRowData_All2015
regularSeason_2015_KNN[1] <- NULL
regularSeason_2015_KNN[2] <- NULL
regularSeason_2015_KNN[3] <- NULL
regularSeason_2015_KNN[4] <- NULL
regularSeason_2015_KNN[3] <- NULL
regularSeason_2015_KNN[2] <- NULL
postSeason_2015_KNN <- singleRowData_All2015_PostSeason
postSeason_2015_KNN[1] <- NULL
postSeason_2015_KNN[2] <- NULL
postSeason_2015_KNN[3] <- NULL
postSeason_2015_KNN[4] <- NULL
postSeason_2015_KNN[3] <- NULL
postSeason_2015_KNN[2] <- NULL
regularSeason_2015_normalized <- as.data.frame(lapply(regularSeason_2015_KNN[1:ncol(regularSeason_2015_KNN)], normalize))
postSeason_2015_normalized <- as.data.frame(lapply(postSeason_2015_KNN[1:ncol(postSeason_2015_KNN)], normalize))

regularSeason_2015_labels <- regularSeason_2015_KNN[1:nrow(regularSeason_2015_KNN), 1]
postSeason_2015_labels <- postSeason_2015_KNN[1:nrow(postSeason_2015_KNN), 1]
knn_prediction <- knn(train = regularSeason_2015_normalized, test = postSeason_2015_normalized, cl = regularSeason_2015_labels, k=4)

```
###Neural Network
```{r}
neuralNet_model <- neuralnet(Win ~ TeamFGM + TeamFGA + TeamFGM3 + TeamFGA3 + TeamFTM + TeamFTA + TeamOR + TeamDR + TeamAST + TeamTO + TeamSTL + TeamBLK + TeamPF, data = regularSeason_2015_normalized)
plot(neuralNet_model)
```

##Testing
```
The premise of March Madness is that an individual has to fill out their bracket before the NCAA Basketball Tournament begins. As a result, when this moment occurs an individual only has the regular season data to made their decisions on, and so I wanted my model to reflect such information limitations that one would face in reality. Therefore, I will use my models above that were trained on Regular Season data and try to predict the winners of March Madness.
```
###Decision Tree Testing
```{r}
singleRowData_All2015_PostSeason_DecisionTree <- singleRowData_All2015_PostSeason
singleRowData_All2015_PostSeason_DecisionTree[1] <- NULL #Remove 'Season' since given
singleRowData_All2015_PostSeason_DecisionTree[3] <- NULL #Remove 'Score'
singleRowData_All2015_PostSeason_DecisionTree[4] <- NULL #Remove 'Score'
win_prediction <- predict(decisionTree_Model, singleRowData_All2015_PostSeason_DecisionTree)

CrossTable(singleRowData_All2015_PostSeason_DecisionTree$Win, win_prediction, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual win', 'predicted win'))
```
```
Correctly predicted 84.3% of last year's March Madness games.
```
###KNN Testing

```{r}
CrossTable(x = postSeason_2015_labels, y = knn_prediction, prop.chisq=FALSE, dnn = c('actual win', 'predicted win'))
```
```
It says that the KNN Model correctly predicted 100% of last year's March Madness games. It seems highly unprobable, and I couldn't find the bug that was causing this.
```

###Neural Network Testing
```{r}
neuralNet_results <- compute(neuralNet_model, postSeason_2015_normalized[1:13])
predicted_strength <- neuralNet_results$net.result
cor(predicted_strength, postSeason_2015_normalized$Win)

new_Pred <- round(predicted_strength, digits=0)
correct <- new_Pred == postSeason_2015_normalized$Win
prop.table(table(correct))
```
According to these test results, the Neural Network correctly predicted 43% of the games.
```


